EoS-insensitive relations after GW170817

### 2019-01-20 paper/bin-love-after-gw170817.tex
minore comments

{\ny{Just some general comments so that we don't forget:
\begin{itemize}
\item The EoS-insensitive relations are called ``binary Love'', ``I-Love-Q'', ``C-Love'' and ``R-Love.'' There is no need to call them ``binary Love EoS-insensitive relations.'' That's too verbose and not what they are called in the literature. I am ok with describing these relations as EoS-insensitive instead of approximately universal, but let's not change the names. 
\item There is some serious tidying up that needs to be done to all of the text. Specifically, there are missing prepositions, like ``the'', in lots of places that makes the reading a bit choppy and we can improve. Zack, when you have a chance, please do a detail read through and fix the flow as much as you can. 
\item Eventually, we need to discuss the abstract and the intro in more detail. 
\item Let's not forget to check the acknowledgements for funding sources. 
\item Let's do a spell check. 
\item Let's remove all commented out text.  
\end{itemize}
}}
The dependence of pressure on density for cold supranuclear matter found inside neutron stars (NSs) remains to be one of the largest mysteries in both nuclear physics and astrophysics.
This internal structure of such stars, known as the equation of state (EoS), is exceedingly important, as it is the determining factor for many NS observables - such as the mass, the radius, and many more.
Unfortunately, terrestrial experiments can only study the EoS to around the nuclear saturation density ($\rho_{\text{sat}} \approx 2.5 \times 10^{14} \text{ g/cm}^3$)~\cite{Li:HeavyIon,Tsang:SymmetryEnergy,Centelles:NeutronSkin,Li:CrossSections,Chen:SymEnergy} (although some temperature-dependent heavy-ion collision experiments may probe higher densities \zc{Katerina do you know of any papers on this?} {\ny{Look up the PREX and CREX experiments at JLab (a good author is Horowitz, e.g. Abrahamyan:2012gp, Horowitz:2012tj, Horowitz:2013wha)}}), making NSs ideal laboratories for constraining ultra dense nuclear matter.

@@

We also present new R-Love relations displaying the relationship between NS radius and tidal deformability, showing maximum EoS variability of $18.88\%$ from the fit discussed in Sec.~\ref{sec:clove}.

@@

\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{EoSs.eps}
\end{center}
\caption{
Small representative sets of the unconstrained (dotted) and constrained (solid) EoSs used in this analysis. 
The constrained EoSs were populated by randomly selecting 100 posterior samples obtained from the GW170817 90\% credible level shown in cyan~\cite{LIGO:posterior}.
Additionally, 100 unconstrained EoSs were generated by randomly sampling spectral coefficients of physically valid EoSs, with no further constrained applied.
The following constraints were also applied to both sets of EoSs: (i) the EoS must have a causal structure within 10\%, and (ii) the EoS must support maximum NS mass of at least $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations.
Finally, the high density core EoSs generated above were stitched to the low-density crust EoS of SLy~\cite{Douchin:2001sv} at $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.
}
\label{fig:eos}
\end{figure}

@@

In this section, we investigate the residuals in $\lambda_0(\Lambda_a,\Lambda_s,q)$ in order to marginalize over the intrinsic error in the binary Love relations.
In particular, we restrict our focus to only the constrained set of EoSs.
Residuals in $\lambda_0$ are computed as $\lambda_0^{\text{fit}}-\lambda_0^{\text{true}}$, where $\lambda_0^{\text{true}}=\lambda_0(\Lambda_a,\Lambda_s,q)$ corresponds to the true value predicted by the sample EoSs, and $\lambda_0^{\text{fit}}=\lambda_0(\Lambda_a^{\text{fit}}(\Lambda_s,q),\Lambda_s,q)$ corresponds to the value obtained through the EoS-insensitive relations found in Sec.~\ref{sec:binLove-nuclear}.
Following Ref.~\cite{Katerina:residuals}, we assume the residuals in $\lambda_0$ observe a Gaussian distribution with mean and standard deviation given by:
\begin{align}
\mu_{\lambda_0}(\Lambda_s,q) &=\frac{\mu_{\Lambda_s}(\Lambda_s)+\mu_{q}(q)}{2},\\ 
\sigma_{\lambda_0} &=\sqrt{\sigma_{\Lambda_s}^2(\Lambda_s) + \sigma_{q}^2(q)}. 
\end{align}
Similar to Ref.~\cite{Katerina:residuals}, we fit the individual components to be:
\begin{align}
\mu_{\Lambda_s}(x) &= \mu_1 x + \mu_2, \label{eq:margFit1}\\ 
\mu_{q}(x) &= \mu_3 x^2 + \mu_4 x + \mu_5, \label{eq:margFit2}\\ 
\sigma_{\Lambda_s}(x) &= \sigma_1 x^{5/2} + \sigma_2 x^{3/2} +  \sigma_3 x^{1/2} + \sigma_4, \label{eq:margFit3}\\ 
\sigma_{q}(x) &= \sigma_5 x^3 + \sigma_6 x^2 + \sigma_7 x + \sigma_8. \label{eq:margFit4}
\end{align}
The fitting parameters $\mu_i$ and $\sigma_i$ are tabulated in Table~\ref{tab:marginalized}.
Figure~\ref{fig:qLsResiduals} shows the $\lambda_0$ residuals binned in both q and $\Lambda_s$, highlighting the error weights across the $(q,\Lambda_s)$ parameter space.
Observe how $q\sim0.5$ and $\Lambda_s\sim2000$ observe the highest errors, while the low and high values of $q$ and $\Lambda_s$ see minimal uncertainties.

@@

>>>>

{\ny{Just some general comments so that we don't forget:
\begin{itemize}
\item The EoS-insensitive relations are called ``binary Love'', ``I-Love-Q'', ``C-Love'' and ``R-Love.'' There is no need to call them ``binary Love EoS-insensitive relations.'' That's too verbose and not what they are called in the literature. I am ok with describing these relations as EoS-insensitive instead of approximately universal, but let's not change the names. 
\item There is some serious tidying up that needs to be done to all of the text. Specifically, there are missing prepositions, like ``the'', in lots of places that makes the reading a bit choppy and we can improve. Zack, when you have a chance, please do a detail read through and fix the flow as much as you can. 
\item Eventually, we need to discuss the abstract and the intro in more detail. 
\item Let's not forget to check the acknowledgements for funding sources. 
\item Let's do a spell check. 
\item Let's remove all commented out text.  
\end{itemize}
}}
The dependence of pressure on density for cold supranuclear matter found inside neutron stars (NSs) remains to be one of the largest mysteries in both nuclear physics and astrophysics.
This internal structure of such stars, known as the equation of state (EoS), is exceedingly important, as it is the determining factor for many NS observables - such as the mass, the radius, and many more.
Unfortunately, terrestrial experiments can only study the EoS to around the nuclear saturation density ($\rho_{\text{sat}} \approx 2.5 \times 10^{14} \text{ g/cm}^3$)~\cite{Li:HeavyIon,Tsang:SymmetryEnergy,Centelles:NeutronSkin,Li:CrossSections,Chen:SymEnergy} (although some temperature-dependent heavy-ion collision experiments may probe higher densities \zc{Katerina do you know of any papers on this?} {\ny{Look up the PREX and CREX experiments at JLab (a good author is Horowitz, e.g. Abrahamyan:2012gp, Horowitz:2012tj, Horowitz:2013wha)}} \kc{these are neutron skin measurement papers, not heavy ion collisions. look at~\cite{Danielewicz:2002pu} }), making NSs ideal laboratories for constraining ultra dense nuclear matter.

@@

We also present new R-Love relations displaying the relationship between NS radius and tidal deformability, showing maximum EoS variability of $18.88\%$\kc{This needs to be expressed in terms of error in the radius, ie. in km} from the fit discussed in Sec.~\ref{sec:clove}.

@@

\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{EoSs.eps}
\end{center}
\caption{
Small representative sets of the unconstrained (dotted) and constrained (solid) EoSs used in this analysis. 
The constrained EoSs were populated by randomly selecting 100 posterior samples obtained from the GW170817 90\% credible level shown in cyan~\cite{LIGO:posterior}.
Additionally, 100 unconstrained EoSs were generated by randomly sampling spectral coefficients of physically valid EoSs, with no further constrained applied.
The following constraints were also applied to both sets of EoSs: (i) the EoS must have a causal structure within 10\%, and (ii) the EoS must support maximum NS mass of at least $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations.
Finally, the high density core EoSs generated above were stitched to the low-density crust EoS of SLy~\cite{Douchin:2001sv} at $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.
\kc{The axis should stop where the 170817 data stops}}
\label{fig:eos}
\end{figure}

@@

\kc{Wait wait wait. Why $\lambda_0$? When we do PE, we do it in terms of $\Lambda_1$ and $\Lambda_2$. So it is these residuals (or equivalently $\Lambda_a$) we need in order to marginalize.}
In this section, we investigate the residuals in $\lambda_0(\Lambda_a,\Lambda_s,q)$ in order to marginalize over the intrinsic error in the binary Love relations.
In particular, we restrict our focus to only the constrained set of EoSs.
Residuals in $\lambda_0$ are computed as $\lambda_0^{\text{fit}}-\lambda_0^{\text{true}}$, where $\lambda_0^{\text{true}}=\lambda_0(\Lambda_a,\Lambda_s,q)$ corresponds to the true value predicted by the sample EoSs, and $\lambda_0^{\text{fit}}=\lambda_0(\Lambda_a^{\text{fit}}(\Lambda_s,q),\Lambda_s,q)$ corresponds to the value obtained through the EoS-insensitive relations found in Sec.~\ref{sec:binLove-nuclear}.
Following Ref.~\cite{Katerina:residuals}, we assume the residuals in $\lambda_0$ observe a Gaussian distribution with mean and standard deviation given by:
\begin{align}
\mu_{\lambda_0}(\Lambda_s,q) &=\frac{\mu_{\Lambda_s}(\Lambda_s)+\mu_{q}(q)}{2},\\ 
\sigma_{\lambda_0} &=\sqrt{\sigma_{\Lambda_s}^2(\Lambda_s) + \sigma_{q}^2(q)}. 
\end{align}
Similar to Ref.~\cite{Katerina:residuals}, we fit the individual components to be:
\begin{align}
\mu_{\Lambda_s}(x) &= \mu_1 x + \mu_2, \label{eq:margFit1}\\ 
\mu_{q}(x) &= \mu_3 x^2 + \mu_4 x + \mu_5, \label{eq:margFit2}\\ 
\sigma_{\Lambda_s}(x) &= \sigma_1 x^{5/2} + \sigma_2 x^{3/2} +  \sigma_3 x^{1/2} + \sigma_4, \label{eq:margFit3}\\ 
\sigma_{q}(x) &= \sigma_5 x^3 + \sigma_6 x^2 + \sigma_7 x + \sigma_8. \label{eq:margFit4}
\end{align}
The fitting parameters $\mu_i$ and $\sigma_i$ are tabulated in Table~\ref{tab:marginalized}.
Figure~\ref{fig:qLsResiduals} shows the $\lambda_0$ residuals binned in both q and $\Lambda_s$, highlighting the error weights across the $(q,\Lambda_s)$ parameter space.
Observe how $q\sim0.5$ and $\Lambda_s\sim2000$ observe the highest errors, while the low and high values of $q$ and $\Lambda_s$ see minimal uncertainties.

@@

\kc{me too}
